---
title: "Data Cleaning and Data Preparation - Phase 2"
output: html_document
Author: AKJ
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let's read the previously cleaned CSV File into a seperate data frame to proceed with the cleaning process. 

```{r Read the prev CSV into a df}
air_quality1 <- read.csv(file='cleaned2_AQ.csv')
```

Let's check if all of our variables are cleaned and does not contain any missing values or data that might mislead our analysis.

```{r checking NA}
is.na.data.frame(air_quality1)
```

So, looks like our dataframe does not have any NA values. But since it is a huge dataframe let's try to visualize if we have any missing data pattern just to be double sure. 

```{r plotting missing data if any}
library(VIM)
explore_missingdata <- aggr(air_quality1, col=c('yellow','red'), numbers=TRUE, sortVars=TRUE, labels=names(data),cex.axis=.4, gap=3, ylab=c("Histogram of missing data","Pattern"))
```

Perfect! Looks like we do not have any missing data values any more. 

```{r dataframe structure}
str(cleaned2_AQ)
```

Finally let's check the summary of our dataframe

```{r}
summary(cleaned2_AQ)
```

Okay, so looks like we are ready for our first level of EDA!  

**Conclusion**
As it is evident from our data frame that for the variable *method_name* we still have some NA values, but since it is a character string hence, it skipped from our identification of missing data values.We shall just leave these NA as is, and make sure that we minimize their impact on our analysis further. Moreover since these are character values or strings and hence statistical imputations such as Mean or Median won't make sense for obvious reasons. 
