---
title: 'Data Cleaning and Data Preparation - Phase 1'
output: html_document
authors: "Ankita Guha and Kara Marsh"
date: "April 22, 2018"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Reading the data file to our dataframe for cleaning the unwanted data. 
```{r read df}
air_quality <- read.csv(file='data/epa_air_quality_annual_summary.csv')
```

Let's try to check our dataframe's consistency with terms of the available 
data.

```{r summary}
# Checking Data Quality
summary(air_quality)
```

It looks like there is a lot of data that is redundant and that won't effect 
our analysis. We will remove those columns to make our analysis easier.

```{r removing redundant variables}
cleaned1_AQ = subset(air_quality, 
                     select = -c(primary_exceedance_count, 
                                 secondary_exceedance_count, 
                                 first_max_non_overlapping_value, 
                                 second_max_non_overlapping_value, 
                                 first_no_max_datetime, 
                                 second_no_max_datetime, 
                                 datum, 
                                 local_site_name, 
                                 city_name, 
                                 cbsa_name, 
                                 first_max_datetime, 
                                 second_max_datetime, 
                                 third_max_datetime, 
                                 fourth_max_datetime, 
                                 poc, 
                                 null_data_count, 
                                 state_code, 
                                 county_code, 
                                 site_num, 
                                 pollutant_standard))
```

Let's try to remove the NA's now

```{r removing some NA}
cleaned2_AQ <- na.omit(cleaned1_AQ)
```

We clean up some of the RAM memory for better data cleaning

```{r clearing memory space}
rm(cleaned1_AQ)
```

For the next phase of cleaning and preparing the data, let's save the 
dataframe into a new CSV File, so that we can continue our analysis with the 
next cleaned dataset in our next RMD File. 

```{r write to a CSV File for analysis in next RMD File}
# Write CSV in R
write.csv(cleaned2_AQ, file = "data/cleaned2_AQ.csv")
```

Let's check the structure of the various variables used in the data frame.

```{r assessing the dataframe structure}
str(cleaned2_AQ)
```
Let's see if we have removed the NA's or not by rechecking our dataframes 
once again. 

```{r summary of dataframe}
summary(cleaned2_AQ)
```

Yep, looks like we have removed almost all the NA's from our dataframe. 



