---
title: "Data Cleaning and Data Preparation - Phase 2"
output: html_document
author: "Ankita Guha, Kara Marsh"
date: "April 22, 2018"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let's read the previously cleaned CSV File into a seperate data frame to proceed with the cleaning process. So the datafile from Kaggle, was blank in method_name. So let's try to put "NA" inside the dataframe and then try to see if we can replace those NA's with something relevant. 

```{r Read the prev CSV into a df}
air_quality1 <- read.csv(file='cleaned2_AQ.csv', na.strings = c("", "NA"))
```

Let's check if all of our variables are cleaned and does not contain any missing values or data that might mislead our analysis.

```{r checking NA}
is.na.data.frame(air_quality1)
```

So, looks like our dataframe does have NA values for method_name. But since it is a huge dataframe let's try to visualize if we have any missing data pattern just to be double sure. 

```{r}
sapply(air_quality1, function(x) sum(is.na(x)))
```

```{r}
make.true.NA <- function(x) if(is.character(x)||is.factor(x)){
                                  is.na(x) <- x %in% c("NA", "N/A","???"); x} else {
                                  x}
```



```{r plotting missing data if any}
library(VIM)
explore_missingdata <- aggr(air_quality1, col=c('lightgreen','red'), numbers=TRUE, sortVars=TRUE, labels=names(data),cex.axis=.4, gap=3, ylab=c("Histogram of missing data","Pattern"))
```
So, we have missing data for method_name. Well, precisely it was blank data and it was not even NA, so we had to figure a way to actually put NA values inside the dataframe and then proceed with our analysis. 


```{r dataframe structure}
#str(air_quality1)
```

Let's clear some of our data frame

```{r}
rm(explore_missingdata, make.true.NA)
```

Okay, so after doing a little EDA we realized that this missing data has a lot of infleunce in our overall dataframe. So we decided to replace them with an easily identifiable variable content name as "Missing Data from Kaggle" so as to make it easy to spot on. 
```{r}
air_quality1$method_name <- as.character(air_quality1$method_name)
 air_quality1$method_name <- ifelse(is.na(air_quality1$method_name), 
             'Missing Data From Kaggle', air_quality1$method_name)
 
```


Finally let's check the summary of our dataframe, if we got any NA value or not!

```{r}
sapply(air_quality1, function(x) sum(is.na(x)))
```

Let's write our cleaned file now for next level of analysis

```{r}
write.csv(air_quality1, file = "cleaned_AQ.csv")
```

Let's free some of our dataframe

```{r}
rm(air_quality1)
```

Okay, so looks like we are ready for our first level of EDA!  

