---
title: "TriCountyEDA"
author: "Kara Marsh"
date: "April 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(ggplot2)
library(plyr)
library(dplyr)
```
## Tri County Analysis

Our first step is to import the data into our environment. 
```{r Reading in Data}
eda <- read.table("cleaned2_AQ.csv", header = TRUE, sep = ",")
```

After importing the data, we filter only the values measuring Ozone, and then 
we find the mean of the arithmetic_mean values. From there, we normalize the 
arithmetic_means.
```{r Ozone filter and normalization}
eda_ozone <- eda %>% filter(parameter_name == 'Ozone')
avg_mean <- mean(eda_ozone$arithmetic_mean)
eda_ozone$norm_mean <- eda_ozone$arithmetic_mean/avg_mean
```

To narrow our analysis, we filter first by state and then find the counties 
which have the greatest number of data points.   
```{r State and County Filters}
mi_ozone <- eda_ozone %>% 
  filter(state_name == "Michigan")

list <- mi_ozone %>% 
  group_by(county_name) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n)) 
```
We create another data frame that just has the three counties with the greatest
number of data points.
```{r}
tricounty_ozone <- mi_ozone %>% 
  filter(county_name == list$county_name[1] | 
           county_name == list$county_name[2] | 
           county_name == list$county_name[3])
```

A basic plot of means over time does not show any useful correlation. 
```{r Plotting Means by Year}
tricounty_ozone %>%
  ggplot() +
  geom_point(aes(x = year, y = norm_mean))  
```
Color coding the values by county also does not let us see any correlation.
```{r}
tricounty_ozone %>%
  ggplot() +
  geom_point(aes(x = year, y = norm_mean, color = county_name))  
```
In the next graph, we color code the metric used to see if that pulls out any 
difference. 
```{r}
tricounty_ozone %>%
  ggplot() +
  geom_point(aes(x = year, y = norm_mean, color = metric_used))  
```
Metric used does seem to correlate to different types of data. Specifically the
red was associated with the daily maxima rather than the daily maximum of an 
average. We will remove the daily maxima from our evalation. 
```{r}
odd_metric = "Daily maxima of observed hourly values (between 9:00 AM and 8:00 PM)"

tricounty_ozone %>%
  filter(metric_used != odd_metric) %>%
  ggplot() +
  geom_point(aes(x = year, y = norm_mean, color = metric_used))  

```
Since those values blend together, we will assume they are roughly equivalent. 
```{r}
tricounty_ozone %>%
  filter(metric_used != odd_metric) %>%
  ggplot() +
  geom_point(aes(x = year, y = norm_mean, color = county_name))  
```

It appears in the graph above that there are multiple values for one county 
in the same year. 
```{r}
tricounty_ozone %>%
  filter(metric_used != odd_metric) %>%
  ggplot() +
  geom_point(aes(x = year, y = norm_mean, color = address)) +
  facet_wrap( ~ county_name)

```



